% ************************** Thesis Acknowledgements *****************************

\begin{acknowledgements}      

My deepest gratitude goes to my supervisor, Prof.~Kenichi Soga, for his 
generous help, support, guidance and encouragement throughout my PhD research. 
I feel privileged to have worked with him; thank you so much for a very 
inspiring and an interesting four years of research.

I would like to express my sincere thanks to my advisor Prof.~Malcolm Bolton, 
for his valuable advice and comments following my first year report.

I am extremely grateful to Prof.~Jean-Yves Delenne, Universit\'{e} Montpellier 
2, France for his support and guidance in developing coupled Lattice Boltzmann 
- Discrete Element Model. I owe a debt of gratitude to Prof.~Farhang Radjai, 
Universit\'{e} Montpellier 2, France for stimulating discussions and guidance 
for this work. I would like to thank Patrick Mutabaruka, Universit\'{e} 
Montpellier 2, France for the collaborative research work and for his support 
during my stay in Montpellier. I owe a debt of gratitude to Samila Bandara, 
EPFL for her continuous support, both as a friend and as a work colleague, for 
her assistance in bringing me up to speed with Material Point Method and for 
her continuous support and discussions.

I would like to express my sincere gratitude to Prof Boominathan, Indian 
Institute of Technology, Madras for his continuous support throughout my 
research career. I sincerely thank my professors at Indian Institute of 
Technology Madras and at Thiagarajar College of Engineering for their support, 
encouragement and for their prior teaching. I would like to thank Dr Vidya 
Bhushan Maji for all his support and for his constant encouragement to pursue 
my PhD at the University of Cambridge, during my research at IITM. A very 
special thanks to Dr.~Nagan, Dr.~Sivagurunathan, and Dr.~Sugunaseelan for their 
constant support and motivation throughout my engineering career. I'm greatly 
indebted to them for my research life.

I owe a very special thanks to my aunt Chitra for all her support, care and 
affection she has showered on me and for always being there for me. I can never 
forget these four years in my life, thank you for making it so special and 
wonderful. Also an extra special thanks to her for getting me started on my 
computing journey. A special thanks to my uncle Sankar and my cousin Karthik 
for their encouragement and making these four years a very memorable experience.

Coming from India, particularly as this was my first time outside the country, 
my friends have always been around to make these four years a wonderful 
experience and making my stay at Cambridge feel like home. A very special 
thanks to Alex Ridge for being such a great friend 
and for being there for me pretty much every single day in the four years of my
PhD. I would like to thank Adam Reid for all his support in easing me in to 
this new country and introducing me to table football. A special thanks to Adam 
for staying with me and supporting me even beyond the visiting hours of the 
hospital after my surgery. A sincere thanks to my extra special friend Nicole 
Grunstra for all our wonderful time together. Extra special thanks to Steven 
Kaneti, who has always been there for me and for all the interesting 
discussions about computing. A special thanks to my friend Charlotte Houldcroft 
for being such a great friend, and for proof-reading the chapters of my thesis. 
My sincere thanks to my friend Katie Fitzpatrick for proof reading the chapters 
of my thesis. I owe an extra special thanks to my friends Renate Fellinger,
Wis Wang-Koh, Anne Musset, James Taylor, Max Hewkin-Smith, Marilena Frisone, 
Auer\'{e}lie Petiot, Isabell Frank, Sarah Williams, Anna Bachmann, Stewart 
Forshaw and all the others, for making my time so far at Cambridge such an 
enjoyable one!

A special thanks again to Alex Ridge, Nicole Grunstra and Steven Kaneti for 
spending sleepless nights to help me format the proceedings of the 
International Symposium on Geomechanics from Micro to Macro. 

I would like to extend my sincere thanks to my friends Karthik, 
Sundaralingam, Meenakshi Sundaram, SME Karthik, Sonu Thomas, Varun Thangamani, 
Anjana Bhashi,  Selvaganesh and Aruna Prasath who have always been there for 
me and for their continuous support and encouragement.

My time at the geotechnical research group would not have been so enjoyable 
without my friends and colleagues. Thank you Tina Schwamb, Denis Garber,
Francesca Burali, Samila Bandara, Ouge Wang, Kirk Ellison, Paul 
Shepley, and Ermao Xu. I am very grateful for the memorable time, 
for the support and encouragement from all of you. A special thanks to the 
Div-D computer officer, Tim Ablett for all his support.

Words cannot express the inspiration, understanding, patience and moral support 
of my parents Soundararajan and Meenakshi, my sister Shenbagam and my 
grandparents, not only in completing my PhD research work but also in all 
aspects of my life. Above all I thank the Gods for their abundant blessings 
showered upon me throughout my life.

Furthermore, I would like to sincerely acknowledge the support of Cambridge 
Commonwealth, European and International Trust for my personal funding during 
this PhD, without which this research would not have been possible. I would 
like to acknowledge the Lundgren Research Award for their financial support 
during the final year of my PhD.

\end{acknowledgements}
